---
title: "Assignment"
author: "Lukas Hofmann"
date: "2024-04-18"
output: html_document
---

# Arnaud Leparmentier’s data

> There is currently a strong controversy as to whether Europe is currently losing ground to the U.S. and if so, how much that is the case. You can see this at the beginning of this article by Arnaud Leparmentier: “In 2008, the eurozone and the US had equivalent gross domestic products (GDP) at current prices of \$14.2 trillion and \$14.8 trillion respectively (€13.1 trillion and €13.6 trillion). Fifteen years on, the eurozone’s GDP is just over \$15 trillion, while US GDP has soared to \$26.9 trillion.”

[Link to the Article](https://www.lemonde.fr/en/opinion/article/2023/09/04/the-gdp-gap-between-europe-and-the-united-states-is-now-80_6123491_23.html)

## 0 - Functions

::: panel-tabset
### General

```{r global_options}
figurenumber <- 1
library(tidyverse)
library(scales)
library(WDI)
library(cowplot)
cEurozone5 <- c("FR", "DE", "GR","IT", "ES")
cEurozone <- c("AT", "BE", "CY", "EE", 
                    "FI", "FR", "DE", "GR", 
                    "IE", "IT", "LV", "LT", 
                    "LU", "MT", "NL", "PT", 
                    "SK", "SI", "ES")
cEurozoneUS <- c("US", "XC")
cEUUS<- c("US", "EU")
countryAll <- c(cEurozoneUS, cEurozone)

```

### GDP function

```{r GdpTab}
gdp <- function(title, indi, countries, bIndex = FALSE, indexindicator = NULL, bCapita = FALSE, capitaindicator = NULL, bUnemployment) {
  data <- WDI(country =countries, indicator = indi) %>%
    mutate(date = paste0(year, "-01-01") %>% as.Date()) %>%
    filter(year >= 2008) %>%
    filter(year <= 2022) %>%
    group_by(country) %>%
    rename(indicatorGDP = indi)
  

  if(!is.null(capitaindicator)) {
    gdp <- data %>%
      select(country,year,date,iso2c, indicatorGDP)
    
    capita <- WDI(country = countries, indicator = capitaindicator) %>%
      filter(year >= 2008) %>%
      filter(year <= 2022) %>%
      rename(indicatorCap = capitaindicator) %>%
      mutate(date = as.Date(paste0(year, "-01-01"))) %>%
      select(country,year,date,iso2c, indicatorCap)
    data <- merge(gdp, capita, by = c("country", "year","date", "iso2c")) %>%
      group_by(country)
  }
  if (bIndex) {
    if (!length(indexindicator)) {
      if(bCapita) {
        data <- data %>%
          mutate(gdpcap = indicatorGDP/indicatorCap) %>%
          mutate(indice = round((100 * gdpcap / gdpcap[year == 2008]), digits = 2))} 
      else{
        data <- data %>%
          mutate(indice = round((100 * indicatorGDP / indicatorGDP[year == 2008]), digits = 2))}
    } else {
      if(bCapita) {
        data <- ungroup(data)
        data <- data %>%
          mutate(gdpcap = indicatorGDP/indicatorCap) %>%
          mutate(x = gdpcap[year == 2008 & iso2c == indexindicator]) %>%
          group_by(country) %>%
          mutate(indice = round((100 * gdpcap / x), digits = 2))} 
      else{
        data <- ungroup(data)
        data <- data %>%
          mutate(x = indicatorGDP[year == 2008 & iso2c == indexindicator]) %>%
          group_by(country) %>%
          mutate(indice = round((100 * indicatorGDP / x), digits = 2))}
    }
  } else {
    if(bCapita) {
      data <- data %>%
          mutate(gdpcap = indicatorGDP/indicatorCap) %>%
          mutate(indice = round(gdpcap, digits = 2))
    }
    else{
      data <- data %>%
        mutate(indice = round(indicatorGDP / 1000000000000,digits = 2))
    }
  }
  
  
  title <- paste("Figure", figurenumber, ":", title)
  graph <- data %>%
    ggplot() + geom_line(aes(x = date, y = indice, color = country)) +
    theme_minimal() + 
    labs(x = "", 
         y = if (bIndex) "2008 = 100" else if (bCapita) "Per Capita" else "in Trillions", 
         title =title, 
         subtitle = paste0("In 2022: ", paste(unique(data$country), "(", data$indice[data$year == 2022], ") ", collapse = ", "),
                     "\nIn 2008: ", paste(unique(data$country), "(", data$indice[data$year == 2008], ") ", collapse = ", "))) +
    #scale_color_manual(values = c("#003399", "#BF0A30")) +
    scale_x_date(breaks = seq(2008, 2100,  2) %>% paste0("-01-01") %>% as.Date,
                 labels = date_format("%Y")) +
    ggtitle(title) +
    theme(legend.position = c(0.2, 0.8),
          legend.title = element_blank())
  
  graph
}
```

### Unemployment function

```{r UnemploymentTab}
unemployment <- function(title, indiGDP, countries, indiUnemployment, capitaindicator=NULL) {
  data <- WDI(country =countries, indicator = indiGDP) %>%
    mutate(date = paste0(year, "-01-01") %>% as.Date()) %>%
    filter(year >= 2008) %>%
    filter(year <= 2022) %>%
    group_by(country) %>%
    rename(indicatorGDP = indiGDP)
  if(!is.null(capitaindicator)) {
    gdp <- data %>%
      select(country,year,date,iso2c,iso3c, indicatorGDP)
    capita <- WDI(country = countries, indicator = capitaindicator) %>%
      filter(year >= 2008) %>%
      filter(year <= 2022) %>%
      rename(indicatorCap = capitaindicator) %>%
      mutate(date = as.Date(paste0(year, "-01-01"))) %>%
      select(country,year,date,iso2c,iso3c, indicatorCap)
    data <- merge(gdp, capita, by = c("country", "year","date", "iso2c","iso3c"))
    data <- data %>%
      group_by(country) %>%
      mutate(indicatorGDP = indicatorGDP/indicatorCap) %>%
      mutate(indice = round(indicatorGDP, digits = 2)) %>%
      select(country,year,date,iso2c,iso3c, indicatorGDP)
   
    } else{
      data <- data %>%
        mutate(indicatorGDP = round(indicatorGDP / 1000000000000,digits = 2)) 
    }
  
  pop <- WDI(country=countries, indicator = indiUnemployment) %>%
    filter(year >= 2008) %>%
    filter(year <= 2022) %>%
    mutate(date = paste0(year, "-01-01") %>% as.Date()) %>%
    group_by(country) %>%
    rename(indicatorCap = indiUnemployment)
  data <- merge(data, pop, by = c("country", "year","date", "iso2c","iso3c")) %>%
      group_by(country) 

  correlations <- data %>%
    summarise(correlation = round(cor(indicatorGDP, indicatorCap), digits = 2))
  
  subtitle <- paste("Correlation for", paste(unique(correlations$country), ":", correlations$correlation, collapse = ", "))
  title <- paste("Figure", figurenumber, ":", title)
  graph <- data %>%
    ggplot() +
    geom_point(aes(x = indicatorGDP, y = indicatorCap, color = country)) +
    geom_smooth(aes(x = indicatorGDP, y = indicatorCap, color = country), method = "lm", se = FALSE) + 

    labs(title = title,
         x = if(is.null(capitaindicator))"GDP in Trillion (Total Value)" else "GDP per Capita",
         y = "Unemployment (%)",
         subtitle = subtitle)+
    theme_minimal()+
    facet_wrap(~country, scales = "free")

  graph
}
```

### Population function

```{r PopulationTab}
population <- function(title, indi, countries, bIndex = FALSE, indexindicator = NULL, bUnemployment = FALSE) {
  pop <- WDI(country=countries, indicator = indi) %>%
    filter(year >= 2008) %>%
    filter(year <= 2022) %>%
    mutate(date = paste0(year, "-01-01") %>% as.Date()) %>%
    group_by(country) %>%
    rename(indicatorCap = indi)
  if (bIndex) {
    if (!length(indexindicator)) {
      pop <- pop %>%
        mutate(indice = round((100 * indicatorCap / indicatorCap[year == 2008]), digits = 2))
    } else {
      pop <- ungroup(pop)
      pop <- pop %>%
        mutate(x = indicatorCap[year == 2008 & iso2c == indexindicator]) %>%
        group_by(country) %>%
        mutate(indice = round((100 * indicatorCap / x), digits = 2))
    }
  } else {
    if (bUnemployment) {
      pop <- pop %>%
        mutate(indice = round(indicatorCap,digits = 2))
    } else{
      pop <- pop %>%
        mutate(indice = round(indicatorCap / 1000000,digits = 2))
    }
  }
  
  
  title <- paste("Figure", figurenumber, ":", title)
  graph <- pop %>%
    ggplot() + geom_line(aes(x = date, y = indice, color = country)) +
    theme_minimal() + 
    labs(x = "", 
         y = if (bIndex) "2008 = 100" else if(bUnemployment) "in percent" else"in Million", 
         title =title, 
         subtitle = paste0("In 2022: ", paste(unique(pop$country), "(", pop$indice[pop$year == 2022], ") ", collapse = ", "),
                     "\nIn 2008: ", paste(unique(pop$country), "(", pop$indice[pop$year == 2008], ") ", collapse = ", "))) +
    scale_x_date(breaks = seq(2008, 2100,  2) %>% paste0("-01-01") %>% as.Date,
                 labels = date_format("%Y")) +
    ggtitle(title) +
    theme(legend.position = c(0.2, 0.8),
          legend.title = element_blank())
  graph
}
```
:::

## 1 - Replicating

```{r}
# GDP current US$
indi = "NY.GDP.MKTP.CD"
#In Values
index = FALSE
title ="GDP (current US$) in total values"
gdp(title, indi, cEurozoneUS, index)
figurenumber <<- figurenumber + 1
#Index of 2008
index = TRUE
title <- "GDP (current US$) with index at 2008"
gdp(title, indi, cEurozoneUS, index)
figurenumber <<- figurenumber + 1
# Europe as an Index
title <- "GDP (current US$) with index at Eurozone in 2008"
gdp(title, indi, cEurozoneUS, index, "XC")
figurenumber <<- figurenumber + 1
```

## 2 -Extending

```{r}
# Unemployment, total (% of total labor force)
indi ="SL.UEM.TOTL.NE.ZS"
title <- "Unemployment in %"
population(title, indi, cEurozoneUS, bIndex = FALSE, bUnemployment = TRUE) 
figurenumber <<- figurenumber + 1
title <- "Unemployement with index 2008"
population(title, indi, cEurozoneUS, bIndex = TRUE,bUnemployment = TRUE)
figurenumber <<- figurenumber + 1
title <- "Unemployement with index 2008+Eurozone"
population(title, indi, cEurozoneUS, bIndex = TRUE, indexindicator="XC",bUnemployment = TRUE)
figurenumber <<- figurenumber + 1

#Population
indi ="SP.POP.TOTL"
title <- "Population in total values"
population(title, indi, cEurozoneUS, bIndex = FALSE) 
figurenumber <<- figurenumber + 1
title <- "Population with index 2008"
population(title, indi, cEurozoneUS, bIndex = TRUE)
figurenumber <<- figurenumber + 1
title <- "Population with index 2008+Eurozone"
population(title, indi, cEurozoneUS, bIndex = TRUE, indexindicator="XC")
figurenumber <<- figurenumber + 1

#Labor force
indi ="SL.TLF.TOTL.IN"
title <- "Labor force in total values"
population(title, indi, cEurozoneUS, bIndex = FALSE) 
figurenumber <<- figurenumber + 1
title <- "Labor force with index 2008"
population(title, indi, cEurozoneUS, bIndex = TRUE)
figurenumber <<- figurenumber + 1
title <- "Labor force with index 2008+Eurozone"
population(title, indi, cEurozoneUS, bIndex = TRUE, indexindicator="XC")
figurenumber <<- figurenumber + 1
```


```{r}
#Labor force and GDP
indi <- "NY.GDP.MKTP.CD"
capita <- "SL.TLF.TOTL.IN"

title <- "GDP / laborforce in total values"
gdp(title, indi, cEurozoneUS, bIndex=FALSE, bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1

title <- "GDP / laborforce - Index 2008"
gdp(title, indi, cEurozoneUS, bIndex=TRUE, bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1

title <- "GDP / laborforce - Index 2008+Eurozone"
gdp(title, indi, cEurozoneUS, bIndex=TRUE, indexindicator="XC", bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1


#Population and GDP 
indi <- "NY.GDP.MKTP.CD"
capita <- "SP.POP.TOTL"

title <- "GDP / population in total values"
gdp(title, indi, cEurozoneUS, bIndex=FALSE, bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1

title <- "GDP / population - Index 2008"
gdp(title, indi, cEurozoneUS, bIndex=TRUE, bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1

title <- "GDP / population - Index 2008+Eurozone"
gdp(title, indi, cEurozoneUS, bIndex=TRUE, indexindicator="XC",bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1

#People in Employment and GDP 
indi <- "NY.GDP.MKTP.CD"
capita <- "SL.EMP.TOTL"

title <- "GDP / People in Employment in total values"
gdp(title, indi, cEurozoneUS, bIndex=FALSE, bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1

title <- "GDP / People in Employment  - Index 2008"
gdp(title, indi, cEurozoneUS, bIndex=TRUE, bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1

title <- "GDP / People in Employment  - Index 2008+Eurozone"
gdp(title, indi, cEurozoneUS, bIndex=TRUE, indexindicator="XC",bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1


# Unemployment and GDP
indiGDP <- "NY.GDP.MKTP.CD"
indiUnemployment ="SL.UEM.TOTL.NE.ZS"
title <- "Unemployment and GDP"
unemployment(title, indiGDP = indiGDP, cEurozoneUS,indiUnemployment = indiUnemployment)
```

## 3 - Robustness


When trying to do comparisons between countries, the [OECD](https://www.oecd.org/fr/sdd/prix-ppa/purchasingpowerparities-frequentlyaskedquestionsfaqs.htm) and [eurostat](https://ec.europa.eu/eurostat/statistics-explained/index.php?title=Beginners:GDP_-_Comparing_GDP:_growth_rate_and_per_capita) recommend a GDP measurement which

1.  takes PPP into account,
2.  consider the size of the population,
3.  and lastly to use a constant dollar value.


```{r}
# 1. PPP
#Total Value
title ="GDP, PPP (current international $)"
indi = "NY.GDP.MKTP.PP.CD"
gdp(title, indi, cEurozoneUS)
figurenumber <<- figurenumber + 1
#Index 2008
index = TRUE
title <- "GDP, PPP (current US$) with index 2008"
gdp(title, indi, cEurozoneUS, index)
figurenumber <<- figurenumber + 1
# Europe 2008 as an Index
title <- "GDP,PPP (current US$)  with index Eurozone 2008"
gdp(title, indi, cEurozoneUS, index, "XC")
figurenumber <<- figurenumber + 1

# 2. Size of the Population
#Total Value
title ="GDP, PPP (current international $) / population"
indi = "NY.GDP.MKTP.PP.CD"
capita <- "SP.POP.TOTL"
gdp(title, indi, cEurozoneUS, bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1
#Index 2008
index = TRUE
title <- "GDP, PPP (current US$) / population with index 2008"
gdp(title, indi, cEurozoneUS, index, bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1
# Europe 2008 as an Index
title <- "GDP. PPP (current US$) / population with index Eurozone 2008"
gdp(title, indi, cEurozoneUS, index, "XC", bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1

#3. Constant Dollar / Without Inflation
#Total Value
title ="GDP, PPP (constant international $) / population"
indi = "NY.GDP.MKTP.PP.KD"
capita <- "SP.POP.TOTL"
gdp(title, indi, cEurozoneUS, bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1
#Index 2008
index = TRUE
title <- "GDP, PPP (constant US$) / population with index 2008"
gdp(title, indi, cEurozoneUS, index, bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1
# Europe 2008 as an Index
title <- "GDP. PPP (constant US$) / population with index Eurozone 2008"
gdp(title, indi, cEurozoneUS, index, "XC", bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1
```

```{r}
# Unemployment and GDP

indiGDP <- "NY.GDP.MKTP.PP.CD"
indiUnemployment ="SL.UEM.TOTL.NE.ZS"
title <- "Unemployment and GDP current, ppp"
unemployment(title, indiGDP = indiGDP, cEurozoneUS,indiUnemployment = indiUnemployment)

indiGDP <- "NY.GDP.MKTP.PP.KD"
indiUnemployment ="SL.UEM.TOTL.NE.ZS"
title <- "Unemployment and GDP constant, ppp"
unemployment(title, indiGDP = indiGDP, cEurozoneUS,indiUnemployment = indiUnemployment)

indiGDP <- "NY.GDP.MKTP.PP.KD"
indiUnemployment ="SL.UEM.TOTL.NE.ZS"
title <- "Unemployment and GDP constant, ppp / population"
unemployment(title, indiGDP = indiGDP, cEurozoneUS,indiUnemployment = indiUnemployment, capitaindicator ="SP.POP.TOTL")
```

## 4 - Breaking it down


```{r}
# Unemployment, total (% of total labor force)
indi ="SL.UEM.TOTL.NE.ZS"
title <- "Unemployment in %"
population(title, indi, cEurozone, bIndex = FALSE, bUnemployment = TRUE) 

indi ="SL.UEM.TOTL.NE.ZS"
title <- "Unemployment in %"
population(title, indi, cEurozone5, bIndex = FALSE, bUnemployment = TRUE) 
figurenumber <<- figurenumber + 1
title <- "Unemployment with index 2008"
population(title, indi, cEurozone5, bIndex = TRUE,bUnemployment = TRUE)
figurenumber <<- figurenumber + 1

indiGDP <- "NY.GDP.MKTP.PP.KD"
indiUnemployment ="SL.UEM.TOTL.NE.ZS"
title <- "Unemployment and GDP constant, ppp / population"
unemployment(title, indiGDP = indiGDP, cEurozone5,indiUnemployment = indiUnemployment, capitaindicator ="SP.POP.TOTL")

#3. Constant Dollar / Without Inflation
#Total Value
title ="GDP, PPP (constant international $) / population"
indi = "NY.GDP.MKTP.PP.KD"
capita <- "SP.POP.TOTL"
gdp(title, indi, cEurozone5, bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1
#Index 2008
index = TRUE
title <- "GDP, PPP (constant US$) / population with index 2008"
gdp(title, indi, cEurozone5, index, bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1
```

## 5 - Writing a short piece


```{r}
# Unemployment, total (% of total labor force)
indi ="SL.UEM.TOTL.NE.ZS"
title <- "Unemployment in %"
population(title, indi, cEUUS, bIndex = FALSE, bUnemployment = TRUE) 
figurenumber <<- figurenumber + 1
title <- "Unemployment with index 2008"
population(title, indi, cEUUS, bIndex = TRUE,bUnemployment = TRUE)
figurenumber <<- figurenumber + 1

indiGDP <- "NY.GDP.MKTP.PP.KD"
indiUnemployment ="SL.UEM.TOTL.NE.ZS"
title <- "Unemployment and GDP constant, ppp / population"
unemployment(title, indiGDP = indiGDP, cEUUS,indiUnemployment = indiUnemployment, capitaindicator ="SP.POP.TOTL")

#3. Constant Dollar / Without Inflation
#Total Value
title ="GDP, PPP (constant international $) / population"
indi = "NY.GDP.MKTP.PP.KD"
capita <- "SP.POP.TOTL"
gdp(title, indi, cEUUS, bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1
#Index 2008
index = TRUE
title <- "GDP, PPP (constant US$) / population with index 2008"
gdp(title, indi, cEUUS, index, bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1
# Europe 2008 as an Index
title <- "GDP. PPP (constant US$) / population with index EU 2008"
gdp(title, indi, cEUUS, index, "EU", bCapita=TRUE, capitaindicator=capita)
figurenumber <<- figurenumber + 1
```
